# Data preparation

This chapter documents how the nestling-level dataset was prepared before model fitting. Each step below gives the purpose of the operation, a foldable code block showing the exact procedure, and the resulting dataset state.

## Source data

The nestling-level workflow starts from one raw workbook, `CF_exp_all_2026.xlsx`, using the sheet `CF_exp.nestlings_analysis`. Only variables used later in the analysis are retained at import.

```{r}
#| label: prep-source-data-code
#| eval: false
#| code-fold: true
#| code-summary: "Show source-data code"

raw_nestlings <- read_excel(
  "CF_exp_all_2026.xlsx",
  sheet = "CF_exp.nestlings_analysis",
  na = c("NA", "", "na", "NaN")
)

analysis_vars <- c(
  "F_RING", "LD", "INC", "CS", "BS", "HD", "SEX", "YEAR",
  "EXP.INC", "EXP.NEST",
  "D2_MASS", "D8_MASS", "D12_MASS", "D12_RTARS", "Early.g", "Late.g", "Growth",
  "LOGMEAN_INCUBATION", "LOGMEAN_NESTLING", "D12_SURV", "GROUP", "BOX", "TEMP.BOX",
  "TEMP.BOX.ENV", "BLOCK"
)

nestlings_selected <- raw_nestlings |>
  dplyr::select(dplyr::all_of(analysis_vars))
```

Result: the source sheet contains 934 nestling records before filtering.

## Variable types

Continuous measurement variables are converted to numeric values. Treatment, identity, year, sex, survival, box, and block variables are converted to factors.

```{r}
#| label: prep-variable-types-code
#| eval: false
#| code-fold: true
#| code-summary: "Show variable-type code"

numeric_vars <- c(
  "LD", "INC", "CS", "BS", "HD", "Growth", "Early.g", "Late.g",
  "D2_MASS", "D8_MASS", "D12_RTARS", "D12_MASS",
  "LOGMEAN_INCUBATION", "LOGMEAN_NESTLING", "TEMP.BOX", "TEMP.BOX.ENV"
)

factor_vars <- c(
  "F_RING", "SEX", "YEAR", "EXP.INC", "EXP.NEST", "D12_SURV", "GROUP", "BOX", "BLOCK"
)

nestlings_typed <- nestlings_selected |>
  dplyr::mutate(
    dplyr::across(dplyr::any_of(c("SEX", "EXP.INC", "EXP.NEST", "GROUP")),
                  ~ stringr::str_squish(as.character(.x))),
    dplyr::across(dplyr::any_of(numeric_vars), ~ suppressWarnings(as.numeric(.x))),
    dplyr::across(dplyr::any_of(factor_vars), as.factor)
  )
```

Result: variables have the classes required by the mixed models.

## Sex filter

The analysis is restricted to nestlings sexed as `F` or `M`. Records with missing or other sex codes are removed before model fitting.

```{r}
#| label: prep-sex-filter-code
#| eval: false
#| code-fold: true
#| code-summary: "Show sex-filter code"

nestlings_typed <- nestlings_typed |>
  dplyr::filter(SEX %in% c("F", "M")) |>
  dplyr::mutate(SEX = droplevels(SEX))
```

Result: 885 sexed nestling records are retained.

## Hatchability dataset

Hatchability is prepared from the brood-level sheet `CF_exp.broods_analysis`. For each brood, clutch size (`CS`) gives the number of eggs and brood size (`BS`) gives the number of hatched young. The brood-level records are expanded to an egg-level binary dataset, where `HATCHED = 1` identifies hatched eggs and `HATCHED = 0` identifies eggs that did not hatch.

```{r}
#| label: prep-hatchability-code
#| eval: false
#| code-fold: true
#| code-summary: "Show hatchability-data code"

raw_broods <- read_excel(
  "CF_exp_all_2026.xlsx",
  sheet = "CF_exp.broods_analysis",
  na = c("NA", "", "na", "NaN")
)

hatchability_broods <- raw_broods |>
  dplyr::select(F_RING, YEAR, BOX, LD, EXP.INC, GROUP, CS, BS) |>
  dplyr::mutate(
    dplyr::across(dplyr::any_of(c("F_RING", "YEAR", "BOX", "EXP.INC", "GROUP")),
                  ~ stringr::str_squish(as.character(.x))),
    dplyr::across(dplyr::any_of(c("LD", "CS", "BS")), ~ suppressWarnings(as.numeric(.x)))
  ) |>
  dplyr::filter(!is.na(F_RING), !is.na(YEAR), !is.na(EXP.INC),
                !is.na(CS), !is.na(BS), CS > 0, BS >= 0, BS <= CS) |>
  dplyr::mutate(
    YEAR = factor(YEAR),
    F_RING = factor(F_RING),
    EXP.INC = factor(EXP.INC),
    GROUP = factor(GROUP),
    CS_sc = as.numeric(scale(CS)),
    unhatched = CS - BS,
    brood_id = paste(F_RING, YEAR, BOX, LD, sep = "_")
  )

hatchability_eggs <- hatchability_broods |>
  tidyr::uncount(weights = CS, .id = "egg_index") |>
  dplyr::group_by(brood_id) |>
  dplyr::mutate(HATCHED = as.integer(dplyr::row_number() <= dplyr::first(BS))) |>
  dplyr::ungroup()
```

Result: the hatchability dataset contains 1170 egg-level records from 178 broods.

## Standardised predictors

Brood size, hatching date, and measured temperatures are standardised before modelling. The original temperature columns `LOGMEAN_INCUBATION` and `LOGMEAN_NESTLING` are used to create clearer scaled-variable names; the final analysis file keeps the scaled variables used by the models.

```{r}
#| label: prep-standardised-predictors-code
#| eval: false
#| code-fold: true
#| code-summary: "Show standardisation code"

prepared_nestlings <- prepared_nestlings |>
  dplyr::mutate(
    HD_sc = as.numeric(scale(HD)),
    BS_sc = as.numeric(scale(BS)),
    incubation_temperature_sc = as.numeric(scale(LOGMEAN_INCUBATION)),
    nestling_temperature_sc = as.numeric(scale(LOGMEAN_NESTLING))
  )
```

Result: the model covariates are `BS_sc`, `HD_sc`, `incubation_temperature_sc`, and `nestling_temperature_sc`.

## Collinearity check

Before fitting the ambient-temperature interaction models, we checked whether the two measured temperature predictors were strongly correlated and whether the fitted models showed problematic collinearity.

The Pearson correlation between `incubation_temperature_sc` and `nestling_temperature_sc` is r = -0.062, based on n = 842 complete records. This indicates no strong linear correlation between the two measured ambient-temperature variables.

The largest VIF in the ambient-temperature interaction models is `4.3`. The highest values occur for the interaction term, which is expected in interaction models. Values remain below the common cautionary threshold of 5, so there is no evidence of severe collinearity.

## Scaled mass index

Day-12 scaled mass index is calculated from day-12 body mass and day-12 tarsus length:

\[
SMI = M_i \left(\frac{L_0}{L_i}\right)^b
\]

where `M_i` is day-12 mass, `L_i` is day-12 tarsus length, `L_0` is the mean tarsus length, and `b` is the standard-major-axis slope approximated as the ratio of standard deviations of log mass and log tarsus.

```{r}
#| label: prep-smi-code
#| eval: false
#| code-fold: true
#| code-summary: "Show SMI code"

smi_base <- prepared_nestlings |>
  tidyr::drop_na(D12_MASS, D12_RTARS) |>
  dplyr::filter(D12_MASS > 0, D12_RTARS > 0) |>
  dplyr::mutate(
    log_mass_n = log(D12_MASS),
    log_tars_n = log(D12_RTARS)
  )

b_sma_chicks <- sd(smi_base$log_mass_n, na.rm = TRUE) / sd(smi_base$log_tars_n, na.rm = TRUE)
L0_chicks <- mean(smi_base$D12_RTARS, na.rm = TRUE)

prepared_nestlings <- prepared_nestlings |>
  dplyr::mutate(
    D12_SMI = dplyr::if_else(
      !is.na(D12_MASS) & !is.na(D12_RTARS) & D12_MASS > 0 & D12_RTARS > 0,
      D12_MASS * (L0_chicks / D12_RTARS)^b_sma_chicks,
      NA_real_
    )
  )
```

Result: day-12 scaled mass index is available for 679 nestlings.

## Treatment counts

The retained data are summarised by the four treatment combinations. Both the number of retained nests and the number of retained nestling records are shown.

```{r}
#| label: prep-treatment-counts-code
#| eval: false
#| code-fold: true
#| code-summary: "Show treatment-count code"

treatment_counts <- prepared_nestlings |>
  dplyr::count(EXP.INC, EXP.NEST, GROUP, name = "nestling_records") |>
  dplyr::left_join(
    prepared_nestlings |>
      dplyr::distinct(F_RING, YEAR, BOX, LD, EXP.INC, EXP.NEST, GROUP) |>
      dplyr::count(EXP.INC, EXP.NEST, GROUP, name = "nests"),
    by = c("EXP.INC", "EXP.NEST", "GROUP")
  ) |>
  dplyr::select(EXP.INC, EXP.NEST, GROUP, nests, nestling_records)
```

```{r}
#| label: prep-treatment-counts-result
#| echo: false
#| message: false
#| warning: false

library(readr)
library(knitr)

counts_file <- "data/treatment_counts.csv"

if (file.exists(counts_file)) {
  read_csv(counts_file, show_col_types = FALSE) |>
    kable()
}
```

## Final analysis dataset

After the preparation checks and treatment counts are created, the file used for modelling is reduced to the columns needed for the analyses. This prevents the shared analysis file from carrying unused raw or intermediate variables.

```{r}
#| label: prep-final-analysis-columns-code
#| eval: false
#| code-fold: true
#| code-summary: "Show final-column selection code"

analysis_columns <- c(
  "F_RING", "YEAR", "SEX",
  "EXP.INC", "EXP.NEST", "GROUP",
  "D2_MASS", "D8_MASS", "D12_MASS", "D12_RTARS",
  "Early.g", "Late.g", "Growth", "D12_SMI", "D12_SURV",
  "BS_sc", "incubation_temperature_sc", "nestling_temperature_sc"
)

analysis_nestlings <- prepared_nestlings |>
  dplyr::select(dplyr::all_of(analysis_columns))

saveRDS(analysis_nestlings, "data/prepared_nestlings.rds")
readr::write_csv(analysis_nestlings, "data/prepared_nestlings.csv")
```

Result: the final analysis dataset contains 885 nestling records and 18 columns.

