# Model strategy

The analyses are organised around two biological questions that are kept separate throughout the website.

First, we ask whether the experimental warming treatment itself affects nestling performance. These are the primary results. They include treatment predictors, sex, brood size, female identity, and year.

Second, we ask whether measured ambient temperature is associated with the same responses. These models are shown separately because they address a different question: not whether the experimental manipulation worked, but whether natural variation in ambient temperature during incubation and the nestling period is related to offspring phenotype or survival.

## Response variables

The dependent variables are:

- hatchability (`HATCHED`),
- day-2 body mass (`D2_MASS`),
- day-8 body mass (`D8_MASS`),
- day-12 body mass (`D12_MASS`),
- early growth (`Early.g`),
- late growth (`Late.g`),
- whole-period growth (`Growth`),
- day-12 scaled mass index (`D12_SMI`),
- day-12 tarsus length (`D12_RTARS`),
- day-12 survival (`D12_SURV`).

Hatchability is modelled separately because it is an egg-level response that occurs before the nestling-period treatment. Day-2 body mass is also treated separately because it occurs before most of the nestling-period treatment. All later nestling responses can include both incubation-period and nestling-period predictors.

## Distributional strategy

For continuous nestling traits, Gaussian mixed models are used as the primary models. This choice is based on interpretability and on the fact that the response distributions do not show strong departures from normality. Student-t models are still fitted because AIC often favours heavier-tailed error distributions, but they are treated as robustness checks rather than as an automatic replacement for the Gaussian models.

AIC values are reported descriptively in the distributional sensitivity analysis. Effects that are consistent between Gaussian and Student-t models are interpreted as more robust. Effects that appear only under one distributional assumption are treated cautiously.

For hatchability and day-12 survival, the primary models use binomial error distributions.

## Primary experiment models

For hatchability, the primary model tests whether incubation-stage treatment affects the probability that eggs hatch:

```{r}
#| label: hatchability-main-model
#| echo: false
cat("HATCHED ~ EXP.INC + CS_sc + (1|F_RING) + (1|YEAR)")
```

`HATCHED` is coded as an egg-level binary response, where `1` means that the egg hatched and `0` means that it did not hatch. The model controls for clutch size (`CS_sc`) and includes female identity and year as random effects.

For day-2 body mass, the primary model tests incubation-stage treatment only:

```{r}
#| label: d2-main-model
#| echo: false
cat("D2_MASS ~ EXP.INC + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

For later continuous traits, the primary experiment model is:

```{r}
#| label: later-main-model
#| echo: false
cat("response ~ EXP.INC + EXP.NEST + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

For day-12 survival, the same fixed-effect structure is fitted with a binomial error distribution:

```{r}
#| label: survival-main-model
#| echo: false
cat("D12_SURV ~ EXP.INC + EXP.NEST + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

## Experimental interaction models

A separate page tests whether incubation-stage warming changes the response to nestling-stage warming:

```{r}
#| label: experiment-interaction-model
#| echo: false
cat("response ~ EXP.INC * EXP.NEST + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

These models address carry-over effects of the experimental treatment. They are presented separately from the primary additive experiment models so that the main treatment effects and interaction question remain easy to read.

## Ambient-temperature models

Measured ambient temperature is analysed on a separate page using an interaction between incubation-period and nestling-period temperature. These models test whether ambient temperature experienced during one breeding stage has carry-over effects on later responses, and whether the association with temperature during one stage depends on the thermal conditions experienced during the other stage:

```{r}
#| label: temperature-interaction-model
#| echo: false
cat("response ~ EXP.INC + EXP.NEST + incubation_temperature_sc * nestling_temperature_sc + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```
