# Model strategy

The analyses are organised around two biological questions that are kept separate throughout the website.

First, we ask whether the experimental warming treatment itself affects nestling performance. These are the primary results. They include treatment predictors, sex, brood size, female identity, and year.

Second, we ask whether measured external temperature is associated with the same responses. These models are shown separately because they address a different question: not whether the experimental manipulation worked, but whether natural variation in outside temperature during incubation and the nestling period is related to offspring phenotype or survival.

## Response variables

The dependent variables are:

- day-2 body mass (`D2_MASS`),
- day-8 body mass (`D8_MASS`),
- day-12 body mass (`D12_MASS`),
- early growth (`Early.g`),
- late growth (`Late.g`),
- whole-period growth (`Growth`),
- day-12 scaled mass index (`D12_SMI`),
- day-12 tarsus length (`D12_RTARS`),
- day-12 survival (`D12_SURV`).

Day-2 body mass is treated separately because it occurs before most of the nestling-period treatment. All later responses can include both incubation-period and nestling-period predictors.

## Distributional strategy

For continuous nestling traits, Gaussian mixed models are used as the primary models. This choice is based on interpretability and on the fact that the response distributions do not show strong departures from normality. Student-t models are still fitted because AIC often favours heavier-tailed error distributions, but they are treated as robustness checks rather than as an automatic replacement for the Gaussian models.

AIC values are reported descriptively in the distributional sensitivity analysis. Effects that are consistent between Gaussian and Student-t models are interpreted as more robust. Effects that appear only under one distributional assumption are treated cautiously.

For day-12 survival, the primary model uses a binomial error distribution.

## Primary experiment models

For day-2 body mass, the primary model tests incubation-stage treatment only:

```{r}
#| label: d2-main-model
#| echo: false
cat("D2_MASS ~ EXP.INC + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

For later continuous traits, the primary experiment model is:

```{r}
#| label: later-main-model
#| echo: false
cat("response ~ EXP.INC + EXP.NEST + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

For day-12 survival, the same fixed-effect structure is fitted with a binomial error distribution:

```{r}
#| label: survival-main-model
#| echo: false
cat("D12_SURV ~ EXP.INC + EXP.NEST + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

## Experimental interaction models

A separate page tests whether incubation-stage warming changes the response to nestling-stage warming:

```{r}
#| label: experiment-interaction-model
#| echo: false
cat("response ~ EXP.INC * EXP.NEST + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

These models address carry-over effects of the experimental treatment. They are presented separately from the primary additive experiment models so that the main treatment effects and interaction question remain easy to read.

## External-temperature models

Measured external temperature is analysed on a separate page using an interaction between incubation-period and nestling-period temperature:

```{r}
#| label: temperature-interaction-model
#| echo: false
cat("response ~ EXP.INC + EXP.NEST + incubation_temperature_sc * nestling_temperature_sc + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

These models ask whether natural variation in external temperature during one stage modifies the association between temperature in the other stage and the response. They are not the primary experimental test; they are used to separate the biological effect of the manipulation from associations with background thermal conditions.
