# Distributional sensitivity analysis

For continuous nestling traits, the primary models use Gaussian errors. Student-t models are fitted as robustness checks because AIC can favour heavier-tailed distributions even when the observed trait distributions do not strongly depart from normality.

AIC is therefore reported as descriptive model-comparison information, not as the sole criterion for choosing the biological model. The main interpretation remains based on the Gaussian mixed models; Student-t results are used to check whether treatment-effect directions and support are robust to the distributional assumption.

## Sensitivity summary

The table below gives a compact Gaussian versus Student-t comparison for the experimental treatment effects. An asterisk in the direction columns marks effects with `p < 0.05`.

```{r}
#| label: distributional-sensitivity-summary
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)
library(tidyr)

coef_file <- "models/model_coefficients.csv"

if (file.exists(coef_file)) {
  response_display_label <- function(response, fallback = response) {
    labels <- c(
      D2_MASS = "Body mass at day 2 (D2_MASS)",
      D8_MASS = "Body mass at day 8 (D8_MASS)",
      D12_MASS = "Body mass at day 12 (D12_MASS)",
      HATCHED = "Hatchability (HATCHED)",
      Early.g = "Early growth (Early.g)",
      Late.g = "Late growth (Late.g)",
      Growth = "Whole-period growth (Growth)",
      D12_SMI = "Scaled mass index at day 12 (D12_SMI)",
      D12_RTARS = "Tarsus length at day 12 (D12_RTARS)",
      D12_SURV = "Survival to day 12 (D12_SURV)"
    )
    label <- labels[[response]]
    if (is.null(label) || is.na(label)) fallback else label
  }

  direction_label <- function(estimate, p_value) {
    direction <- case_when(
      is.na(estimate) ~ "not estimable",
      estimate < 0 ~ "negative",
      estimate > 0 ~ "positive",
      TRUE ~ "zero"
    )
    support <- case_when(
      is.na(p_value) ~ "",
      p_value < 0.05 ~ "*",
      TRUE ~ ""
    )
    paste0(direction, support)
  }

  read_csv(coef_file, show_col_types = FALSE) |>
    filter(
      !response %in% c("HATCHED", "D12_SURV"),
      model_name %in% c("gaussian_mean", "student_mean"),
      grepl("^EXP\\.", term)
    ) |>
    mutate(
      effect = recode(
        term,
        "EXP.INCHeated" = "Incubation warming",
        "EXP.NESTHeated" = "Nestling-stage warming",
        .default = term
      ),
      direction = direction_label(estimate, p_value),
      supported = !is.na(p_value) & p_value < 0.05
    ) |>
    mutate(response_label = mapply(response_display_label, response, response_label)) |>
    select(response_label, effect, model_name, direction, supported) |>
    pivot_wider(
      names_from = model_name,
      values_from = c(direction, supported),
      names_sep = "_"
    ) |>
    mutate(
      same_direction = gsub("\\*", "", direction_gaussian_mean) == gsub("\\*", "", direction_student_mean),
      same_support = supported_gaussian_mean == supported_student_mean,
      `Main effect robust?` = case_when(
        same_direction & same_support & supported_gaussian_mean ~ "yes",
        same_direction & same_support & !supported_gaussian_mean ~ "yes, direction only",
        same_direction & !same_support ~ "cautious",
        TRUE ~ "no/cautious"
      ),
      Interpretation = case_when(
        `Main effect robust?` == "yes" ~ "robust",
        `Main effect robust?` == "yes, direction only" ~ "same direction, weak statistical support",
        same_direction & !same_support ~ "same direction, support differs",
        TRUE ~ "distribution-sensitive"
      )
    ) |>
    select(
      Response = response_label,
      Effect = effect,
      `Main effect robust?`,
      `Gaussian direction` = direction_gaussian_mean,
      `Student-t direction` = direction_student_mean,
      Interpretation
    ) |>
    arrange(Response, Effect) |>
    kable()
} else {
  cat("Distributional sensitivity summary is not available yet.")
}
```

## Candidate model set

The detailed AIC table is retained as background information. It is not used as an automatic rule for selecting the biological model.

```{r}
#| label: distributional-candidate-models
#| echo: false
#| message: false
#| warning: false

comparison_file <- "models/model_comparisons.csv"

if (file.exists(comparison_file)) {
  response_display_label <- function(response, fallback = response) {
    labels <- c(
      D2_MASS = "Body mass at day 2 (D2_MASS)",
      D8_MASS = "Body mass at day 8 (D8_MASS)",
      D12_MASS = "Body mass at day 12 (D12_MASS)",
      HATCHED = "Hatchability (HATCHED)",
      Early.g = "Early growth (Early.g)",
      Late.g = "Late growth (Late.g)",
      Growth = "Whole-period growth (Growth)",
      D12_SMI = "Scaled mass index at day 12 (D12_SMI)",
      D12_RTARS = "Tarsus length at day 12 (D12_RTARS)",
      D12_SURV = "Survival to day 12 (D12_SURV)"
    )
    label <- labels[[response]]
    if (is.null(label) || is.na(label)) fallback else label
  }

  read_csv(comparison_file, show_col_types = FALSE) |>
    filter(!response %in% c("HATCHED", "D12_SURV")) |>
    mutate(
      response_label = mapply(response_display_label, response, response_label),
      role = if_else(selected_model, "Primary Gaussian model", "Sensitivity model"),
      AIC = round(AIC, 2),
      delta_AIC = round(delta_AIC, 2)
    ) |>
    arrange(response_label, AIC) |>
    select(response_label, model_name, model_family, dispersion_formula, AIC, delta_AIC, role) |>
    kable(col.names = c("Response", "Model", "Family", "Dispersion", "AIC", "Delta AIC", "Role"))
} else {
  cat("Distributional model-comparison table is not available yet.")
}
```

## Interpretation rule

Effects that show the same direction in Gaussian and Student-t models are treated as directionally robust. The asterisk indicates whether the effect also has statistical support (`p < 0.05`). Effects that change direction, or have support only under one distributional assumption, are interpreted cautiously as distribution-sensitive results.

When the estimated effect changed direction between Gaussian and Student-t models, this indicates that the result was sensitive to distributional assumptions. We therefore do not interpret this effect as robust evidence for a biological relationship.

## Dispersion effects

For continuous responses, we also fitted location-scale models in `glmmTMB`, allowing residual dispersion to differ between experimental treatments. These models ask whether the experiment affected not only the mean response, but also the variability of the trait. Dispersion effects are interpreted cautiously and separately from the primary mean effects.

Positive dispersion estimates indicate higher residual variance relative to the reference group, whereas negative estimates indicate lower residual variance. The table below reports dispersion terms for scale models; the column `Best-supported scale model` marks responses for which a scale model had the lowest AIC among the candidate models.

```{r}
#| label: dispersion-effects-table
#| echo: false
#| message: false
#| warning: false

dispersion_file <- "models/dispersion_effects/dispersion_coefficients.csv"

if (file.exists(dispersion_file)) {
  read_csv(dispersion_file, show_col_types = FALSE) |>
    filter(term != "(Intercept)", best_model_has_dispersion) |>
    mutate(
      estimate = round(estimate, 3),
      std_error = round(std_error, 3),
      statistic = round(statistic, 3),
      p_value = round(p_value, 4),
      direction = case_when(
        estimate > 0 ~ "higher variance",
        estimate < 0 ~ "lower variance",
        TRUE ~ "no change"
      )
    ) |>
    select(response_label, model_name, term, direction, estimate, std_error, statistic, p_value) |>
    arrange(response_label, term) |>
    kable(col.names = c("Response", "Scale model", "Dispersion term", "Direction", "Estimate", "SE", "z", "p"))
} else {
  cat("Dispersion-effect estimates are not available yet.")
}
```

## Treatment-group contrasts

As an additional sensitivity analysis, we also compare the four combined treatment groups (`CONCON`, `CONEXP`, `EXPCON`, and `EXPEXP`). These contrasts are descriptive and are used to make group-level differences easier to inspect. They do not replace the primary factorial models, which remain the main biological tests of incubation-stage warming, nestling-stage warming, and their interaction.

For nestling traits, the sensitivity model is:

```text
response ~ GROUP + SEX + BS_sc + (1 | F_RING) + (1 | YEAR)
```

For hatchability, the model is:

```text
HATCHED ~ GROUP + CS_sc + (1 | F_RING) + (1 | YEAR)
```

The hatchability group comparison is included only as a descriptive sensitivity check, because hatchability occurs before the nestling-stage treatment.

```{r}
#| label: treatment-group-estimated-means
#| echo: false
#| message: false
#| warning: false

group_means_file <- "models/group_contrasts/group_estimated_means.csv"

if (file.exists(group_means_file)) {
  read_csv(group_means_file, show_col_types = FALSE) |>
    mutate(
      estimate = if_else(is.na(prob), emmean, prob),
      lower = asymp.LCL,
      upper = asymp.UCL,
      estimate = round(estimate, 3),
      SE = round(SE, 3),
      lower = round(lower, 3),
      upper = round(upper, 3)
    ) |>
    select(response_label, GROUP, n, estimate, SE, lower, upper) |>
    arrange(response_label, GROUP) |>
    kable(col.names = c("Response", "Group", "N", "Estimated mean/probability", "SE", "Lower 95% CI", "Upper 95% CI"))
} else {
  cat("Treatment-group estimated means are not available yet.")
}
```

```{r}
#| label: treatment-group-pairwise-contrasts
#| echo: false
#| message: false
#| warning: false
#| results: asis

group_contrasts_file <- "models/group_contrasts/group_pairwise_contrasts.csv"

if (file.exists(group_contrasts_file)) {
  contrast_table <- read_csv(group_contrasts_file, show_col_types = FALSE) |>
    mutate(
      estimate_display = if_else(is.na(odds.ratio), estimate, odds.ratio),
      scale = if_else(is.na(odds.ratio), "difference", "odds ratio"),
      estimate_display = round(estimate_display, 3),
      SE = round(SE, 3),
      p.value = round(p.value, 4)
    ) |>
    select(response_label, contrast, scale, estimate_display, SE, p.value) |>
    arrange(response_label, p.value)

  cat("<details>\n")
  cat("<summary>Show all pairwise treatment-group contrasts</summary>\n\n")
  print(
    kable(
      contrast_table,
      col.names = c("Response", "Contrast", "Scale", "Estimate", "SE", "Tukey-adjusted p")
    )
  )
  cat("\n</details>\n")
} else {
  cat("Treatment-group pairwise contrasts are not available yet.")
}
```
