# Distributional sensitivity analysis

For continuous nestling traits, the primary models use Gaussian errors. Student-t models are fitted as robustness checks because AIC can favour heavier-tailed distributions even when the observed trait distributions do not strongly depart from normality.

AIC is therefore reported as descriptive model-comparison information, not as the sole criterion for choosing the biological model. The main interpretation remains based on the Gaussian mixed models; Student-t results are used to check whether treatment-effect directions and support are robust to the distributional assumption.

## Sensitivity summary

The table below gives a compact Gaussian versus Student-t comparison for the experimental treatment effects. An asterisk in the direction columns marks effects with `p < 0.05`.

```{r}
#| label: distributional-sensitivity-summary
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)
library(tidyr)

coef_file <- "models/model_coefficients.csv"

if (file.exists(coef_file)) {
  direction_label <- function(estimate, p_value) {
    direction <- case_when(
      is.na(estimate) ~ "not estimable",
      estimate < 0 ~ "negative",
      estimate > 0 ~ "positive",
      TRUE ~ "zero"
    )
    support <- case_when(
      is.na(p_value) ~ "",
      p_value < 0.05 ~ "*",
      TRUE ~ ""
    )
    paste0(direction, support)
  }

  read_csv(coef_file, show_col_types = FALSE) |>
    filter(
      response != "D12_SURV",
      model_name %in% c("gaussian_mean", "student_mean"),
      grepl("^EXP\\.", term)
    ) |>
    mutate(
      effect = recode(
        term,
        "EXP.INCHeated" = "Incubation warming",
        "EXP.NESTHeated" = "Nestling-stage warming",
        .default = term
      ),
      direction = direction_label(estimate, p_value),
      supported = !is.na(p_value) & p_value < 0.05
    ) |>
    select(response_label, effect, model_name, direction, supported) |>
    pivot_wider(
      names_from = model_name,
      values_from = c(direction, supported),
      names_sep = "_"
    ) |>
    mutate(
      same_direction = gsub("\\*", "", direction_gaussian_mean) == gsub("\\*", "", direction_student_mean),
      same_support = supported_gaussian_mean == supported_student_mean,
      `Main effect robust?` = case_when(
        same_direction & same_support & supported_gaussian_mean ~ "yes",
        same_direction & same_support & !supported_gaussian_mean ~ "yes, direction only",
        same_direction & !same_support ~ "cautious",
        TRUE ~ "no/cautious"
      ),
      Interpretation = case_when(
        `Main effect robust?` == "yes" ~ "robust",
        `Main effect robust?` == "yes, direction only" ~ "same direction, weak statistical support",
        same_direction & !same_support ~ "same direction, support differs",
        TRUE ~ "distribution-sensitive"
      )
    ) |>
    select(
      Response = response_label,
      Effect = effect,
      `Main effect robust?`,
      `Gaussian direction` = direction_gaussian_mean,
      `Student-t direction` = direction_student_mean,
      Interpretation
    ) |>
    arrange(Response, Effect) |>
    kable()
} else {
  cat("Distributional sensitivity summary is not available yet.")
}
```

## Candidate model set

The detailed AIC table is retained as background information. It is not used as an automatic rule for selecting the biological model.

```{r}
#| label: distributional-candidate-models
#| echo: false
#| message: false
#| warning: false

comparison_file <- "models/model_comparisons.csv"

if (file.exists(comparison_file)) {
  read_csv(comparison_file, show_col_types = FALSE) |>
    filter(response != "D12_SURV") |>
    mutate(
      role = if_else(selected_model, "Primary Gaussian model", "Sensitivity model"),
      AIC = round(AIC, 2),
      delta_AIC = round(delta_AIC, 2)
    ) |>
    arrange(response_label, AIC) |>
    select(response_label, model_name, model_family, dispersion_formula, AIC, delta_AIC, role) |>
    kable(col.names = c("Response", "Model", "Family", "Dispersion", "AIC", "Delta AIC", "Role"))
} else {
  cat("Distributional model-comparison table is not available yet.")
}
```

## Interpretation rule

Effects that show the same direction in Gaussian and Student-t models are treated as directionally robust. The asterisk indicates whether the effect also has statistical support (`p < 0.05`). Effects that change direction, or have support only under one distributional assumption, are interpreted cautiously as distribution-sensitive results.

When the estimated effect changed direction between Gaussian and Student-t models, this indicates that the result was sensitive to distributional assumptions. We therefore do not interpret this effect as robust evidence for a biological relationship.
