# Sensitivity interaction analysis

The primary models use `GROUP` to represent the four experimental combinations. This is clean for presentation, but it does not directly test whether incubation treatment interacts with the nestling-phase treatment. Therefore, a sensitivity analysis decomposes `GROUP` into its two experimental phases.

## Biological question

This sensitivity analysis asks whether effects associated with incubation and incubation-period ambient temperature depend on conditions during the nestling phase. In other words, it tests whether early thermal conditions and later nestling-phase conditions combine additively or interactively when predicting fitness-related nestling traits.

## Model structure

For later nestling traits, the additive sensitivity model is:

```{r}
#| label: sensitivity-additive-formula
#| echo: false
cat("response ~ EXP.INC + EXP.NEST + incubation_temperature_sc + nestling_temperature_sc + SEX + BLOCK + (1|F_RING) + (1|YEAR)")
```

The interaction sensitivity model is:

```{r}
#| label: sensitivity-interaction-formula
#| echo: false
cat("response ~ EXP.INC * EXP.NEST + incubation_temperature_sc * nestling_temperature_sc + SEX + BLOCK + (1|F_RING) + (1|YEAR)")
```

For continuous traits, both mean-only and location-scale versions are fitted. In the location-scale models, residual dispersion is allowed to vary by the interaction between incubation and nestling treatment:

```{r}
#| label: sensitivity-dispformula
#| echo: false
cat("dispformula = ~ EXP.INC * EXP.NEST")
```

Day-2 mass is not included in this interaction sensitivity set because day-2 mass precedes most of the nestling-phase treatment and therefore is not the right response for testing incubation-by-nestling-phase interactions.

## Candidate-model comparisons

```{r}
#| label: sensitivity-comparison-files
#| echo: false
#| message: false
#| warning: false

library(readr)
library(dplyr)
library(knitr)

comparison_files <- list.files(
  "models/sensitivity_interactions",
  pattern = "^sensitivity_interaction_comparison_.*\\.csv$",
  full.names = TRUE
)

if (length(comparison_files) > 0) {
  comparisons <- lapply(comparison_files, read_csv, show_col_types = FALSE) |>
    bind_rows()

  comparisons |>
    select(any_of(c("response", "candidate_model", "AIC", "AIC_wt", "BIC", "RMSE", "n_observations"))) |>
    kable()
} else {
  cat("Sensitivity interaction comparisons will appear here after running scripts/04_fit_sensitivity_interactions.R.")
}
```

## Model summaries

```{r}
#| label: sensitivity-summary-files
#| echo: false
#| message: false
#| warning: false

summary_files <- list.files(
  "models/sensitivity_interactions",
  pattern = "^summary_.*\\.txt$",
  full.names = TRUE
)

if (length(summary_files) > 0) {
  tibble::tibble(summary_file = summary_files) |>
    kable()
} else {
  cat("Sensitivity model summaries will appear here after running scripts/04_fit_sensitivity_interactions.R.")
}
```