# Results

This chapter presents the primary experiment-only models response by response. These models test whether the warming treatment affected each trait, while keeping measured external temperature out of the primary experimental test. Temperature models are shown on a separate page.

```{r}
#| label: results-setup
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)
library(glue)
library(tidyr)

selected_file <- "models/selected_models.csv"
coef_file <- "models/model_coefficients.csv"
artifact_file <- "models/model_artifacts.csv"

if (file.exists(selected_file) && file.exists(coef_file) && file.exists(artifact_file)) {
  selected_models <- read_csv(selected_file, show_col_types = FALSE)
  model_coefficients <- read_csv(coef_file, show_col_types = FALSE)
  model_artifacts <- read_csv(artifact_file, show_col_types = FALSE)
} else {
  selected_models <- tibble()
  model_coefficients <- tibble()
  model_artifacts <- tibble()
}

distribution_sensitive_terms <- model_coefficients |>
  filter(
    response != "D12_SURV",
    model_name %in% c("gaussian_mean", "student_mean"),
    grepl("^EXP\\.", term),
    !is.na(estimate)
  ) |>
  mutate(direction = sign(estimate)) |>
  select(response, term, model_name, direction) |>
  tidyr::pivot_wider(names_from = model_name, values_from = direction) |>
  filter(!is.na(gaussian_mean), !is.na(student_mean), gaussian_mean != student_mean) |>
  select(response, term)

fmt_num <- function(x, digits = 3) ifelse(is.na(x), NA, formatC(x, digits = digits, format = "f"))
link_path <- function(path, text = basename(path)) {
  if (length(path) == 0 || is.na(path) || !file.exists(path)) return("")
  glue("[{text}]({path})")
}

response_question <- function(response, response_label) {
  questions <- c(
    D2_MASS = "Does incubation-stage warming affect body mass at day 2?",
    D8_MASS = "Do incubation-stage and nestling-stage warming affect body mass at day 8?",
    D12_MASS = "Do incubation-stage and nestling-stage warming affect body mass at day 12?",
    Early.g = "Do incubation-stage and nestling-stage warming affect early nestling growth?",
    Late.g = "Do incubation-stage and nestling-stage warming affect late nestling growth?",
    Growth = "Do incubation-stage and nestling-stage warming affect whole-period growth?",
    D12_SMI = "Do incubation-stage and nestling-stage warming affect day-12 scaled mass index?",
    D12_RTARS = "Do incubation-stage and nestling-stage warming affect day-12 tarsus length?",
    D12_SURV = "Do incubation-stage and nestling-stage warming affect survival to day 12?"
  )
  if (!is.na(questions[[response]])) questions[[response]] else paste("Model for", response_label)
}

print_saved_summary <- function(summary_file) {
  if (length(summary_file) == 0 || is.na(summary_file) || !file.exists(summary_file)) {
    cat("Saved summary file not found.\n\n")
    return(invisible(NULL))
  }
  summary_lines <- readLines(summary_file, warn = FALSE)
  cat("<details>\n")
  cat("<summary>Show saved model summary</summary>\n\n")
  cat("```text\n")
  cat(paste(summary_lines, collapse = "\n"))
  cat("\n```\n")
  cat("</details>\n\n")
}
```

## Response-level results

```{r}
#| label: response-level-results
#| echo: false
#| message: false
#| warning: false
#| results: asis
#| fig-width: 7.2
#| fig-height: 4.8

if (nrow(selected_models) > 0) {
  for (i in seq_len(nrow(selected_models))) {
    response <- selected_models$response[i]
    response_label <- selected_models$response_label[i]
    selected_name <- selected_models$selected_model[i]

    artifact <- model_artifacts |>
      filter(response == !!response, selected_model) |>
      slice(1)

    cat("\n\n### ", response_label, "\n\n", sep = "")

    cat("#### Dependent variable\n\n")
    cat("`", response, "`: ", response_question(response, response_label), "\n\n", sep = "")

    cat("#### Model\n\n")
    cat("**Primary model:** `", selected_name, "`  \n", sep = "")
    cat("**Formula:** `", selected_models$formula[i], "`  \n", sep = "")
    cat("**Sample size:** `", selected_models$n[i], "`  \n", sep = "")
    if (!is.na(selected_models$dispersion_formula[i]) && selected_models$dispersion_formula[i] != "not used") {
      cat("**Dispersion formula:** `", selected_models$dispersion_formula[i], "`  \n", sep = "")
    }
    cat("\n")

    cat("#### Fixed-effect estimates\n\n")
    sensitive_terms <- distribution_sensitive_terms |>
      filter(response == !!response)

    model_coefficients |>
      filter(response == !!response, model_name == !!selected_name) |>
      left_join(
        sensitive_terms |> mutate(distribution_sensitive = TRUE),
        by = c("response", "term")
      ) |>
      mutate(
        distribution_sensitive = if_else(is.na(distribution_sensitive), FALSE, distribution_sensitive),
        term = if_else(distribution_sensitive, paste0(term, "*"), term),
        estimate = fmt_num(estimate),
        std_error = fmt_num(std_error),
        statistic = fmt_num(statistic),
        p_value = fmt_num(p_value, 4)
      ) |>
      select(term, estimate, std_error, statistic, p_value) |>
      kable(col.names = c("Term", "Estimate", "SE", "z", "p")) |>
      print()
    cat("\n\n")

    if (nrow(sensitive_terms) > 0) {
      cat("*The marked experimental effect changed direction between Gaussian and Student-t models. This result is treated as distribution-sensitive and is not interpreted here as robust biological evidence; see the Distribution sensitivity section.*\n\n")
    }

    cat("#### Saved model summary\n\n")
    if (nrow(artifact) == 1) {
      cat("Summary PDF: ", link_path(artifact$summary_pdf[1]), "  \n", sep = "")
      cat("Diagnostic PDF: ", link_path(artifact$diagnostic_pdf[1]), "\n\n", sep = "")
      print_saved_summary(artifact$summary_txt[1])
    } else {
      cat("Saved summary files are not available for this response.\n\n")
    }

    raw_fig <- file.path("figures", "raw_distributions", paste0("raw_", response, ".png"))
    effect_fig <- file.path("figures", "model_effects", paste0("model_effect_", response, ".png"))

    if (file.exists(effect_fig)) {
      cat("#### Model prediction figure\n\n")
      if (response == "D12_SURV") {
        cat("This plot uses model predictions as the main signal. Dark points and lines show predicted survival probability from the primary binomial mixed model, vertical intervals are 95% confidence intervals, and pale bars in the background show observed survival proportions.\n\n")
      } else {
        cat("This plot uses model predictions as the main signal. Dark points and lines show fitted means from the primary mixed model, vertical intervals are 95% confidence intervals, and pale jittered points in the background show the original observations.\n\n")
      }
      cat(glue("![]({effect_fig})"), "\n\n")
    }

    if (file.exists(raw_fig)) {
      cat("#### Raw-data distribution\n\n")
      if (response == "D12_SURV") {
        cat("This raw-data plot shows observed survival proportions within each treatment combination. Bars are proportions of nestlings classified as dead or survived; no model-based uncertainty interval is shown.\n\n")
      } else {
        cat("This raw-data plot shows the response distribution within treatment groups. The box centre line is the median, the box spans the interquartile range (25th-75th percentiles), whiskers show the boxplot range, and points are individual nestling observations.\n\n")
      }
      cat(glue("![]({raw_fig})"), "\n\n")
    }
  }
} else {
  cat("Saved model-result files are not available yet.\n")
}
```
