# Experimental interaction models

This page tests whether the effect of nestling-stage warming depends on incubation-stage warming. Biologically, this asks whether conditions experienced before hatching modify later responses to the nestling-period treatment.

The interaction model is:

```{r}
#| label: experiment-interaction-formula
#| echo: false
cat("response ~ EXP.INC * EXP.NEST + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

Day-2 mass is not included here because it occurs before most of the nestling-stage treatment.

## Interaction terms

```{r}
#| label: experiment-interaction-terms
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)

interaction_file <- "models/experiment_interaction/experiment_interaction_terms.csv"

response_display_label <- function(response, fallback = response) {
  labels <- c(
    D2_MASS = "Body mass at day 2 (D2_MASS)",
    D8_MASS = "Body mass at day 8 (D8_MASS)",
    D12_MASS = "Body mass at day 12 (D12_MASS)",
    Early.g = "Early growth (Early.g)",
    Late.g = "Late growth (Late.g)",
    Growth = "Whole-period growth (Growth)",
    D12_SMI = "Scaled mass index at day 12 (D12_SMI)",
    D12_RTARS = "Tarsus length at day 12 (D12_RTARS)",
    D12_SURV = "Survival to day 12 (D12_SURV)"
  )
  label <- labels[[response]]
  if (is.null(label) || is.na(label)) fallback else label
}

if (file.exists(interaction_file)) {
  read_csv(interaction_file, show_col_types = FALSE) |>
    mutate(response_label = mapply(response_display_label, response, response_label)) |>
    mutate(
      estimate = round(estimate, 3),
      std_error = round(std_error, 3),
      statistic = round(statistic, 3),
      p_value = round(p_value, 4)
    ) |>
    select(response_label, term, estimate, std_error, statistic, p_value) |>
    kable(col.names = c("Response", "Interaction term", "Estimate", "SE", "z", "p"))
} else {
  cat("Experiment-interaction results are not available yet.")
}
```

## Saved model summaries

```{r}
#| label: experiment-interaction-summaries
#| echo: false
#| message: false
#| warning: false
#| results: asis

summary_files <- list.files(
  "models/experiment_interaction",
  pattern = "^summary_.*interaction.*\\.txt$",
  full.names = TRUE
)

if (length(summary_files) > 0) {
  for (summary_file in sort(summary_files)) {
    cat("\n\n### ", tools::file_path_sans_ext(basename(summary_file)), "\n\n", sep = "")
    cat("<details>\n")
    cat("<summary>Show saved model summary</summary>\n\n")
    cat("```text\n")
    cat(paste(readLines(summary_file, warn = FALSE), collapse = "\n"))
    cat("\n```\n")
    cat("</details>\n")
  }
} else {
  cat("Saved experiment-interaction summaries are not available yet.")
}
```
