# Experimental interaction models

We initially tested whether the effect of nestling-stage warming depended on the thermal conditions experienced during incubation. Biologically, this tests whether exposure to warming before hatching modifies the response to warming during the nestling period. In other words, these models ask whether the effect of the nestling-period treatment differs between nests that were heated or not heated during incubation.

For each response variable, we fitted the following interaction model:

```text
response ~ EXP.INC * EXP.NEST + SEX + BS_sc + (1 | F_RING) + (1 | YEAR)
```

where `EXP.INC` indicates whether the nest was heated during incubation, and `EXP.NEST` indicates whether the nest was heated during the nestling period.

Across traits, the interaction between incubation-stage and nestling-stage warming was not statistically supported. We therefore present these models as planned interaction tests, while interpreting the main treatment effects from the primary models reported above.

Day-2 body mass was not included in this set of interaction models because this measurement was taken before most of the nestling-stage treatment had occurred.

## Interaction terms

```{r}
#| label: experiment-interaction-terms
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)

interaction_file <- "models/experiment_interaction/experiment_interaction_terms.csv"

response_display_label <- function(response, fallback = response) {
  labels <- c(
    D2_MASS = "Body mass at day 2 (D2_MASS)",
    D8_MASS = "Body mass at day 8 (D8_MASS)",
    D12_MASS = "Body mass at day 12 (D12_MASS)",
    Early.g = "Early growth (Early.g)",
    Late.g = "Late growth (Late.g)",
    Growth = "Whole-period growth (Growth)",
    D12_SMI = "Scaled mass index at day 12 (D12_SMI)",
    D12_RTARS = "Tarsus length at day 12 (D12_RTARS)",
    D12_SURV = "Survival to day 12 (D12_SURV)"
  )
  label <- labels[[response]]
  if (is.null(label) || is.na(label)) fallback else label
}

if (file.exists(interaction_file)) {
  read_csv(interaction_file, show_col_types = FALSE) |>
    mutate(response_label = mapply(response_display_label, response, response_label)) |>
    mutate(
      estimate = round(estimate, 3),
      std_error = round(std_error, 3),
      statistic = round(statistic, 3),
      p_value = round(p_value, 4)
    ) |>
    select(response_label, term, estimate, std_error, statistic, p_value) |>
    kable(col.names = c("Response", "Interaction term", "Estimate", "SE", "z", "p"))
} else {
  cat("Experiment-interaction results are not available yet.")
}
```

## Saved model summaries

```{r}
#| label: experiment-interaction-summaries
#| echo: false
#| message: false
#| warning: false
#| results: asis

summary_files <- list.files(
  "models/experiment_interaction",
  pattern = "^summary_.*interaction.*\\.txt$",
  full.names = TRUE
)

if (length(summary_files) > 0) {
  for (summary_file in sort(summary_files)) {
    cat("\n\n### ", tools::file_path_sans_ext(basename(summary_file)), "\n\n", sep = "")
    cat("<details>\n")
    cat("<summary>Show saved model summary</summary>\n\n")
    cat("```text\n")
    cat(paste(readLines(summary_file, warn = FALSE), collapse = "\n"))
    cat("\n```\n")
    cat("</details>\n")
  }
} else {
  cat("Saved experiment-interaction summaries are not available yet.")
}
```
