# Ambient temperature models

This page addresses a separate question from the primary experiment results: whether measured ambient temperature is associated with offspring traits and survival.

The aim is not to replace the experimental treatment models, but to show whether natural variation in ambient temperature during incubation and the nestling period is related to the same responses. These models include the treatment predictors and then add the interaction between measured incubation-period and nestling-period ambient temperature:

```{r}
#| label: ambient-temperature-formula
#| echo: false
cat("response ~ EXP.INC + EXP.NEST + incubation_temperature_sc * nestling_temperature_sc + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

Day-2 mass is not included because it occurs before most of the nestling-period treatment.

## Interaction terms

```{r}
#| label: ambient-temperature-interaction-terms
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)
library(glue)

interaction_file <- "models/sensitivity_temperature_interaction/temperature_interaction_terms.csv"

response_display_label <- function(response, fallback = response) {
  labels <- c(
    D2_MASS = "Body mass at day 2 (D2_MASS)",
    D8_MASS = "Body mass at day 8 (D8_MASS)",
    D12_MASS = "Body mass at day 12 (D12_MASS)",
    Early.g = "Early growth (Early.g)",
    Late.g = "Late growth (Late.g)",
    Growth = "Whole-period growth (Growth)",
    D12_SMI = "Scaled mass index at day 12 (D12_SMI)",
    D12_RTARS = "Tarsus length at day 12 (D12_RTARS)",
    D12_SURV = "Survival to day 12 (D12_SURV)"
  )
  label <- labels[[response]]
  if (is.null(label) || is.na(label)) fallback else label
}

if (file.exists(interaction_file)) {
  read_csv(interaction_file, show_col_types = FALSE) |>
    mutate(response_label = mapply(response_display_label, response, response_label)) |>
    mutate(
      estimate = round(estimate, 3),
      std_error = round(std_error, 3),
      statistic = round(statistic, 3),
      p_value = round(p_value, 4)
    ) |>
    select(response_label, term, estimate, std_error, statistic, p_value) |>
    kable(col.names = c("Response", "Interaction term", "Estimate", "SE", "z", "p"))
} else {
  cat("Ambient-temperature interaction results are not available yet.")
}
```

## Response-level models

For each response, the section below shows the fitted ambient-temperature interaction model, the saved model summary, and the corresponding prediction figure. The prediction figure is the main visual interpretation. Panel A shows the incubation-temperature effect at low, mean, and high nestling-stage temperature. Panel B shows the nestling-stage temperature effect at low, mean, and high incubation temperature. Lines show population-level model predictions, shaded ribbons show 95% confidence intervals, and grey points show raw observations.

Prediction figures are shown only for interaction terms with statistical support or near-support (`p <= 0.09`).

```{r}
#| label: ambient-temperature-response-level
#| echo: false
#| message: false
#| warning: false
#| results: asis

response_labels <- c(
  D8_MASS = "Body mass at day 8 (D8_MASS)",
  D12_MASS = "Body mass at day 12 (D12_MASS)",
  Early.g = "Early growth (Early.g)",
  Late.g = "Late growth (Late.g)",
  Growth = "Whole-period growth (Growth)",
  D12_SMI = "Scaled mass index at day 12 (D12_SMI)",
  D12_RTARS = "Tarsus length at day 12 (D12_RTARS)",
  D12_SURV = "Survival to day 12 (D12_SURV)"
)

responses <- names(response_labels)

formula_text <- "response ~ EXP.INC + EXP.NEST + incubation_temperature_sc * nestling_temperature_sc + SEX + BS_sc + (1|F_RING) + (1|YEAR)"

interaction_file <- "models/sensitivity_temperature_interaction/temperature_interaction_terms.csv"
figure_responses <- character(0)

if (file.exists(interaction_file)) {
  figure_responses <- readr::read_csv(interaction_file, show_col_types = FALSE) |>
    dplyr::filter(!is.na(p_value), p_value <= 0.09) |>
    dplyr::pull(response)
}

for (response in responses) {
  response_label <- response_labels[[response]]
  summary_file <- file.path(
    "models",
    "sensitivity_temperature_interaction",
    paste0("summary_", response, ifelse(response == "D12_SURV", "_interaction_binomial.txt", "_interaction_gaussian.txt"))
  )
  figure_file <- file.path(
    "figures",
    "temperature_interactions",
    paste0("temperature_interaction_", response, ".png")
  )

  cat("\n\n### ", response_label, "\n\n", sep = "")
  cat("**Model:** `", sub("^response", response, formula_text), "`\n\n", sep = "")

  cat("#### Saved model summary\n\n")
  if (file.exists(summary_file)) {
    cat("<details>\n")
    cat("<summary>Show saved model summary</summary>\n\n")
    cat("```text\n")
    cat(paste(readLines(summary_file, warn = FALSE), collapse = "\n"))
    cat("\n```\n")
    cat("</details>\n\n")
  } else {
    cat("Saved model summary is not available for this response.\n\n")
  }

  if (response %in% figure_responses) {
    cat("#### Prediction figure\n\n")
  }

  if (file.exists(figure_file) && response %in% figure_responses) {
    cat(glue("![]({figure_file})"), "\n\n")
    cat("Lines show population-level model predictions. In panel A, lines are estimated at low (-1 SD), mean, and high (+1 SD) nestling-stage temperature. In panel B, lines are estimated at low, mean, and high incubation temperature. Shaded ribbons show 95% confidence intervals. Grey points show raw observations.\n\n")
  }
}
```
