# External-temperature models

This page addresses a separate question from the primary experiment results: whether measured external temperature is associated with offspring traits and survival.

The aim is not to replace the experimental treatment models, but to show whether natural variation in external temperature during incubation and the nestling period is related to the same responses. These models include the treatment predictors and then add the interaction between measured incubation-period and nestling-period temperature:

```{r}
#| label: external-temperature-formula
#| echo: false
cat("response ~ EXP.INC + EXP.NEST + incubation_temperature_sc * nestling_temperature_sc + SEX + BS_sc + (1|F_RING) + (1|YEAR)")
```

Day-2 mass is not included because it occurs before most of the nestling-period treatment.

## Collinearity check

Before interpreting the temperature-interaction models, we check whether the two measured temperature predictors are strongly correlated and whether the fitted models show problematic collinearity.

```{r}
#| label: temperature-collinearity
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)

correlation_file <- "models/collinearity/temperature_correlation.csv"
vif_file <- "models/collinearity/temperature_model_vif.csv"

if (file.exists(correlation_file) && file.exists(vif_file)) {
  temperature_correlation <- read_csv(correlation_file, show_col_types = FALSE)
  vif_table <- read_csv(vif_file, show_col_types = FALSE)

  cat(
    "The Pearson correlation between `incubation_temperature_sc` and `nestling_temperature_sc` is `r = ",
    round(temperature_correlation$pearson_r[1], 3),
    "` based on `n = ",
    temperature_correlation$n_complete[1],
    "` complete records. This indicates no strong linear correlation between the two measured temperature variables.\n\n",
    sep = ""
  )

  cat(
    "The largest VIF in the temperature-interaction models is `",
    round(max(vif_table$VIF, na.rm = TRUE), 2),
    "`. The highest values occur for the interaction term, which is expected in interaction models. Values remain below the common cautionary threshold of 5, so there is no evidence of severe collinearity.\n\n",
    sep = ""
  )

  vif_table |>
    filter(grepl("temperature", Term)) |>
    mutate(
      VIF = round(VIF, 2),
      SE_factor = round(SE_factor, 2)
    ) |>
    select(response_label, Term, VIF, SE_factor) |>
    kable(col.names = c("Response", "Term", "VIF", "SE inflation factor"))
} else {
  cat("Collinearity diagnostics are not available yet.")
}
```

## Temperature-interaction terms

```{r}
#| label: temperature-interaction-terms
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(readr)
library(knitr)

interaction_file <- "models/sensitivity_temperature_interaction/temperature_interaction_terms.csv"

if (file.exists(interaction_file)) {
  read_csv(interaction_file, show_col_types = FALSE) |>
    mutate(
      estimate = round(estimate, 3),
      std_error = round(std_error, 3),
      statistic = round(statistic, 3),
      p_value = round(p_value, 4)
    ) |>
    select(response_label, term, estimate, std_error, statistic, p_value) |>
    kable(col.names = c("Response", "Interaction term", "Estimate", "SE", "z", "p"))
} else {
  cat("Temperature-interaction results are not available yet.")
}
```

## Prediction figures

```{r}
#| label: temperature-interaction-figures
#| echo: false
#| message: false
#| warning: false
#| results: asis

library(glue)

figures <- list.files(
  "figures/temperature_interactions",
  pattern = "^temperature_interaction_.*\\.png$",
  full.names = TRUE
)

if (length(figures) > 0) {
  for (fig in sort(figures)) {
    response <- sub("^temperature_interaction_", "", tools::file_path_sans_ext(basename(fig)))
    cat("\n\n### ", response, "\n\n", sep = "")
    cat("Lines show predicted values across incubation-period external temperature at low, average, and high nestling-period external temperature. Ribbons are 95% confidence intervals; pale points show original observations when readable.\n\n")
    cat(glue("![]({fig})"), "\n\n")
  }
} else {
  cat("Temperature-interaction figures are not available yet.")
}
```

## Saved model summaries

```{r}
#| label: temperature-interaction-summaries
#| echo: false
#| message: false
#| warning: false
#| results: asis

summary_files <- list.files(
  "models/sensitivity_temperature_interaction",
  pattern = "^summary_.*interaction.*\\.txt$",
  full.names = TRUE
)

if (length(summary_files) > 0) {
  for (summary_file in sort(summary_files)) {
    cat("\n\n### ", tools::file_path_sans_ext(basename(summary_file)), "\n\n", sep = "")
    cat("<details>\n")
    cat("<summary>Show saved model summary</summary>\n\n")
    cat("```text\n")
    cat(paste(readLines(summary_file, warn = FALSE), collapse = "\n"))
    cat("\n```\n")
    cat("</details>\n")
  }
} else {
  cat("Saved temperature-interaction summaries are not available yet.")
}
```
